---
## title: "Instrumental Variables in Case-Shiller Methodology" 
author: "Your Name" 
date: "`r Sys.Date()`" 
output: html_document
---
# Introduction

The Case-Shiller methodology states that an **instrumental variables estimator** is used:

where:

- &#x20;is the matrix of independent variables (potentially endogenous).
- &#x20;is the vector of observed price changes.
- &#x20;is the instrument matrix, constructed by replacing price levels in  with  or , indicating the sales timing.

The document does not explicitly discuss iteration, but because  contains terms involving , we **must iterate** to find consistent estimates.

# Implementation in R

```{r
knitr::opts_chunk$set(echo = TRUE)
```

## **Simulating Price Data**

```{r}
# Simulated price data (replace with actual Case-Shiller data)
P <- matrix(c(
  100, 90,  80,  # P11, P21, P51
   95, 85,  75,  # P10, P20, P52
  120, 110, 90,  # P32, P42, P52 (time period 2)
  105, 95,  85   # P30, P40, P51
), ncol = 3, byrow = TRUE)
```

## **Setting Up IV Estimation with Iteration**

```{r}
# Initialize beta estimates
beta <- matrix(c(1, 1), ncol = 1)  # Initial guesses for β1 and β2
tolerance <- 1e-6
max_iter <- 100
iter <- 0
converged <- FALSE

# Define the instrument matrix Z (1s and -1s indicating sales)
Z <- matrix(c(1, 1, 1, 1, 1, 1), ncol = 2)  # Example structure

# Iterative estimation
while (iter < max_iter && !converged) {
  iter <- iter + 1

  # Construct the X matrix dynamically based on beta
  X <- matrix(c(
    P[1,1] + P[2,1] + P[3,1],  # Numerator for β1
    P[4,1] + P[5,1] + beta[2,1] * P[6,1],  # Denominator for β1
    P[1,2] + P[2,2] + P[3,2],  # Numerator for β2
    P[4,2] + P[5,2] + beta[1,1] * P[6,2]   # Denominator for β2
  ), ncol = 2, byrow = TRUE)

  # Apply the IV estimator β = (Z'X)^-1 Z'Y
  beta_new <- solve(t(Z) %*% X) %*% (t(Z) %*% P[,3])

  # Check for convergence
  if (max(abs(beta_new - beta)) < tolerance) {
    converged <- TRUE
  }

  beta <- beta_new  # Update beta estimates
}
```

## **Final Results**

```{r}
# Output final estimates
cat("Estimated beta1:", beta[1,1], "\n")
cat("Estimated beta2:", beta[2,1], "\n")
```

# Conclusion

- The **Case-Shiller methodology** states that they use an **instrumental variables estimator**.
- However, because  itself depends on , we must **iterate** to achieve convergence.
- This is **not explicitly mentioned** in the document, making it non-trivial to implement.

This approach generalizes to higher-dimensional problems, but for large-scale datasets (e.g., 235,000 rows, 30 columns), we would use:

1. **Sparse matrices** (`Matrix` package).
2. **Efficient solvers** like `qr.solve()` or `glmnet`.
3. **A dedicated simultaneous equations solver** (`systemfit` package).

Would you like an implementation using `systemfit` for efficiency? 🚀

